Best Practices for Multiple Alchemical Free Energy Calculations [Article v\versionnumber]

### 2025-05-12 paper/manuscript.tex
Adding an illustration of the timing diferences.

%Geometrz
An alternative to SIS-based MCSS methods is geometric bipartite matching for atom mapping, as implemented in tools such as pmx~\cite{gapsys2015pmx} and Kartograf~\cite{ries2024kartograf}. These algorithms enable rapid mappings due to their reduced computational complexity ($O(N) = N^3$), provided that well-aligned input geometries are available~\cite{ries2024kartograf}, through e.g. restraint-docking or prior-alignment. To accommodate topological variations, Kartograf applies a subsequent filtering step following the 3D mapping process.

@@

In terms of computational efficiency, the choice between methods largely depends on input conditions. If molecular structures are pre-aligned, the geometric approach offers a substantial speed advantage. Conversely, when ligands lack predefined 3D coordinates, the SIS-based method provides a easy solution, allowing coordinate generation at a later stage. This nuanced selection process ensures optimal performance based on the specific characteristics of the molecular system under investigation.

>>>>

% Geometry
An alternative to SIS-based MCSS methods is geometric bipartite matching for atom mapping, as implemented in tools such as pmx~\cite{gapsys2015pmx} and Kartograf~\cite{ries2024kartograf}. These algorithms enable rapid mappings due to their reduced computational complexity ($O(N) = N^3$), provided that well-aligned input geometries are available~\cite{ries2024kartograf}, through e.g. restraint-docking or prior-alignment. To accommodate topological variations, Kartograf applies a subsequent filtering step following the 3D mapping process.

@@

In terms of computational efficiency, the choice between methods largely depends on input conditions. If molecular structures are pre-aligned, the geometric approach offers a substantial speed advantage. The magnitude of the speed advantage is dependent on the number of mappings required and the number of atoms to be considered. Here we want to exemplify the effect by using the example of \textit{Ries et al.} calculating all to all mappings for given sets of molecules. The first example was a small set of 14 molecules and an average $38.6$ atoms per molecule; both approaches needed only $3~s$ to calculate the all to all mappings. A more challenging dataset with 16 molecules and an average of $74$ atoms per molecule increased the calculation duration to $26~s$ for the SIS based approach with LOMAP and $8~s$ for the geometric approach with Kartograf. Finally, using the double amount of molecules leads to a significant difference of $58~s$ for the geometric approach and $11~min$ and $16~s$ for the SIS based approach, indicating the algorithmic complexityty difference. The complexity, however, does not only scale with the number of molecules. The last example of \textit{Ries et al.} uses a protein with $9349$ atoms and mutates one residue, leading to two molecules. The problem was solved with the geometric approach in $2~s$. The SIS-based approach was stopped after $4~h$. \cite{ries2024kartograf}

Conversely, when ligands lack predefined 3D coordinates, the SIS-based method provides a easy solution, allowing coordinate generation at a later stage. This nuanced selection process ensures optimal performance based on the specific characteristics of the molecular system under investigation.
